% !TeX root = ../../thesis.tex
\chapter{Anisotropic interface energy}
\section{Capillary vector formalism}
Let the interface energy $\sigma$ depend on the local orientation of the interface described by its unit normal $\bm{n}$, i.e. $\sigma=\sigma(\bm{n})=\sigma_0 f(\bm{n})$. The function $f(\bm{n})$ is called anisotropy function and $\sigma_0$  is a scaling parameter with dimensions of interface energy. Formal requirements on the anisotropy function can be found e.g. in~\cite{Kobayashi2001}. A very useful formalism for thermodynamic description of these anisotropic interfaces is so-called capillary vector $\bm{\xi}$ (a.k.a. Cahn-Hoffmann vector or $\xi$-vector), defined as 
\begin{equation}
     \bm{\xi} = \mathrm{grad}(r\sigma(\bm{n}))\,,
 \end{equation}
where $r$ is radial distance from origin, which can be also seen as magnitude of scaled normal vector $\bm{r}=r\bm{n}$. Level sets of the scalar function $r\sigma(\bm{n})$ are geometrically similar to the anisotropy function $f(\bm{n})$, which is continuously scaled by the radius $r$.

Capillary vector was introduced and investigated by Hoffman and Cahn in~\cite{Hoffman1972,Cahn1974}, who showed that this formalism is consistent with major laws governing behavior of (anisotropic) interfaces like Gibbs-Thompson equation (relating chemical potential and isotropic surface curvature), Herring's equation (chemical potential on anisotropic curved surface), Wulff shape construction, force balance at multi-junctions and others. 

A defining property of capillary vector is that its component normal to the surface $\xi_\perp$ is in magnitude equal to the local value of the interface energy and the surface-tangent component points in the direction of steepest change in interface energy and is proportional to the change. More specifically  
\begin{align}
     \xi_\perp &= \bm{\xi}\cdot\bm{n}=\sigma(\bm{n}) \\
     \xi_\| &= \left(\frac{\partial \sigma}{\partial \theta}\right)_{max} \,,
 \end{align}
where $\theta$ is the angle representing change in orientation of $\bm{n}$ in the direction of maximal angular rate change. The vector can thus be written
\begin{equation}
     \bm{\xi}(\bm{n}) = \sigma(\bm{n})\bm{n} + \left(\frac{\partial \sigma}{\partial \theta}\right)_{max} \bm{t} \,,
 \end{equation}
with $\bm{t}$ being unit vector tangent to the surface pointing in the described direction. Useful is expression of capillary vector in spherical coordinates 
\begin{equation}
     \bm{\xi}(\theta,\phi) = \sigma(\theta,\phi)\bm{\hat{r}} + \left(\frac{\partial \sigma}{\partial \theta}\right)\bm{\hat{\theta}} + \frac{1}{\sin(\theta)}\left(\frac{\partial \sigma}{\partial \phi}\right)\bm{\hat{\phi}} \,,
 \end{equation}
which is only function of the polar and azimuthal angle $\theta,\phi$, respectively. In 2D in polar coordinates the normal and tangent vector can be written as functions of the interface normal angle $\theta$ as
\begin{equation} \label{eq_xivec_2D}
     \bm{\xi}(\theta) = \sigma(\theta)\begin{bmatrix}
           \cos(\theta)  \\
           \sin(\theta) 
      \end{bmatrix} + \left(\frac{\partial \sigma}{\partial \theta}\right)\begin{bmatrix}
           -\sin(\theta) \\
           \cos(\theta)  
      \end{bmatrix} \,.
 \end{equation}

\section{Interface stiffness}

\section{Wulff shape}
Wulff shape is such shape, which minimizes the interface energy given a constant volume. In 2D it can be obtained as the inner convex hull of all tangent straight lines to the anisotropy function $f(\theta)$. This can be used to define the Wulff shape $W$ as a set of points $\bm{x}$
\begin{equation}
     W = \{ \bm{x}\in\mathbb{R}| \bm{x}\cdot\bm{n}\leq \sigma(\bm{n})\} \,,
 \end{equation}
which is a geometrical representation of what was said above - projection of the position vector $\bm{x}$ of a point within the Wulff shape to the interface normal vector $\bm{n}$ must be smaller than the anisotropic interface energy. This defines the inner hull of the tangent lines.

When the heads of capillary vectors for all interface normal angles $\theta$ are connected (having a common starting point in the origin), so-called $\xi$-plot is obtained. It turns out, that the $\xi$-plot coincides with the Wulff shape~\cite{Hoffman1972}. Hence, the expression~\eqref{eq_xivec_2D} can be used to plot the 2D Wulff shape in cartesian coordinates. This is entirely true as long as there are no corners or facets on the Wulff shape, in which cases the use of capillary vecor for Wulff shape plotting must be appropriately modified (see below for details).

In~\cite{Cahn1974} the authors re-state he problem of finding equilibrium shapes of crystals with anisotropic interfaces under general conditions. The shape is found as a solution to equation
\begin{equation} \label{eq_xivec_equilibrium_shape}
     \Delta\Omega_v + \nabla_S\cdot \bm{\xi} = 0 \,,
 \end{equation}
where $\Delta\Omega_v$ is the bulk driving force, specifically the difference in grand-canonical potential of the two phases, and $\nabla_S\cdot$ is surface divergence. The surface $\bm{r}$ is seeked, complying with the above equation. A simple trick is used to find the shape of an isolated particle. Surface divergence over the surface itself is simply $\nabla_S\cdot\bm{r}=2$, hence we can write
\begin{align}
     1 = \nabla_S\cdot\bm{r}/2 &= -\nabla_S\cdot\bm{\xi}/\Delta\Omega_v \\
      \nabla_S\cdot(\bm{r} +2\bm{\xi}/\Delta\Omega_v) &=0 \,,
 \end{align}
leading to the result
\begin{equation}
     \bm{r} = -\frac{2}{\Delta\Omega_v}\bm{\xi} \,.
 \end{equation}

Two different types of singularities may occur in Wulff shapes, which must be treated in a specific way: corners and facets. 

Corners represent a discontinuity of normal angle along the Wulff shape perimeter. They occur when the anisotropy of the interface energy is such, that interface stiffness gets negative for some normal angles $\theta_f$. These angles are called forbidden and they are thermodynamically unstable. For this reason, they are not present on the Wulff shape. Additionally, there are normal angles $\theta$ which are thermmodynamically stable but are not present on the Wulff shape. The latter ones correspond to angles, where the inverse interface energy $1/\sigma(\theta)$ is non-convex but still have positive interface stiffness. All forbidden angles fall into this non-convex region. By replacing the non-convex part of this function by straight lines a regularized inverse anisotropy function $1/\bar{\sigma}(\theta)$ is obtained. Wulff shape corresponding to the regularized anisotropy function $\bar{\sigma}(\theta)$ does not contain the "ears" behind the corners. The interface stiffness of both the allowed and forbidden angles which form the "ears" were thus set to zero.

The facets were not treated in this work, hence they are not further discussed.

\section{Force balance in triple junction and stability of its configuration} \label{sec_intro_trijun_forcebalance_stability}
A triple junction in 2D is simply a point and force balance is achieved if the capillary vectors of the three interfaces sum to zero, i.e. 
\begin{equation} \label{eq_trijun_forcebalance_sum_xivec}
     \bm{\xi}_{12}+\bm{\xi}_{23}+\bm{\xi}_{13}=0
 \end{equation}
The problem of force balance in triple junctions was treated in detail in~\cite{Marks2012}. Even though the capillary vector formalism was not used, a condition equivalent to~\eqref{eq_trijun_forcebalance_sum_xivec} was obtained from geometrical arguments (note that the 90-degree rotation does not change meaning of the equation)
\begin{equation} \label{eq_force_balance_3jun_aniso}
     \sum_{i=1}^3 \left[ \sigma_i(\vartheta_i)\hat{\bm{t}}_i +\frac{\mathrm{d} \sigma_i}{\mathrm{d} \vartheta_i}\hat{\bm{n}}_i \right] = 0 \,,
 \end{equation}
where the subscripts now denote interfaces numbered from 1 to 3. $\vartheta_1,\vartheta_2,\vartheta_3$ are angles under which the tangent of each interface is oriented (all w.r.t.\ x axis),  the anisotorpic interface energies are $\sigma_i(\vartheta_i)$. The triple junction configuration is defined by the triplet of the angles $\vartheta_i$. However, it was pointed out, that the above condition is only necessary for the triple junction configuration to be stable, i.e. it does not guarantee the configuration stability. Validity of~\eqref{eq_force_balance_3jun_aniso} assures a stationary point in Gibbs free energy, but in order to have a stable configuration, the energy must be in local minimum. To confirm that, second derivatives of Gibbs free energy with respect to angles $\vartheta_1,\vartheta_2,\vartheta_3$ must be assessed, giving rise to the following two conditions (both of which must hold in stable configuration)
\begin{align} \label{eq_3jun_aniso_stabcond1}
     0 <& \sum_{i=1}^3 \left[ \frac{\mathrm{d}^2 \sigma_i}{\mathrm{d} \vartheta_i^2} + \sigma_i(\vartheta_i) \right]\frac{\sin^2(\vartheta_i)}{d_i} \\  \label{eq_3jun_aniso_stabcond2}
     0 <& \left[ \frac{\mathrm{d}^2 \sigma_1}{\mathrm{d} \vartheta_1^2} + \sigma_1(\vartheta_1) \right]\left[ \frac{\mathrm{d}^2 \sigma_2}{\mathrm{d} \vartheta_2^2} + \sigma_2(\vartheta_2) \right] \frac{\sin^2(\vartheta_1-\vartheta_2)}{d_1d_2} +    \\
      &\quad +\left[ \frac{\mathrm{d}^2 \sigma_2}{\mathrm{d} \vartheta_2^2} + \sigma_2(\vartheta_2) \right]\left[ \frac{\mathrm{d}^2 \sigma_3}{\mathrm{d} \vartheta_3^2} +  \sigma_3(\vartheta_3) \right] \frac{\sin^2(\vartheta_2-\vartheta_3)}{d_2d_3} + \nonumber \\
      &\quad + \left[ \frac{\mathrm{d}^2 \sigma_1}{\mathrm{d} \vartheta_1^2} + \sigma_1(\vartheta_1) \right]\left[ \frac{\mathrm{d}^2 \sigma_3}{\mathrm{d} \vartheta_3^2} + \sigma_3(\vartheta_3) \right]  \frac{\sin^2(\vartheta_3-\vartheta_1)}{d_3d_1} \nonumber \,,
 \end{align}
where $d_i$ are lengths from the triple junctions to anchor points, which are not relevant for the current work. Each term in the brackets is interface stiffness of the respective interface.

If the second condition is equal to 0, these conditions are not conclusive, i.e. it is not possible to assess the nature of the stationary point (higher-order derivatives are needed then). 

The assessment of stability of triple junction configuration is an important part of discussion of an equilibrium shape of anisotropic particle  on a substrate plane.

\section{Equilibrium shapes of anisotorpic particle on planar substrate}
This problem has attracted considerable attention since it is of high practical importance, especially in applications like nucleation, preparation of functional deposits for electrochemical catalysis or in solid-state dewetting. The classical solution to the problem is by Kaischew for crystalline anisotropy with faceted chape\textit{(ref 1.50 in Milchev), year 1951} and Winterbottom1967 (interface energy as continuous function). Here it is demonstrated using the capillary vector formalism. 

In the following only 2D cross section is discussed. It is assumed that the particle-substrate interface (denoted here with index 1) is linear with the substrate-parent phase (index 3), but with anti-parallel normal vector. Let the substrate plane be parallel with x axis. The particle-parent phase interface is denoted by index 2. The normal vectors of the respective interfaces in the contact point are $\hat{\bm{n}}_1=(0,-1)^{\mathrm{T}}$, $\hat{\bm{n}}_3=(0,1)^{\mathrm{T}}$ and $\hat{\bm{n}}_2$ is to be found from~\eqref{eq_trijun_forcebalance_sum_xivec}. Apparently, with the interfaces 1 and 3 fixed at orientations $\theta_1=-\pi/2$ and $\theta_3=\pi/2$, the force balance can be reduced to single equation, when the particle-parent phase capillary vector $\bm{\xi}_2$ is projected to the surface normal $\hat{\bm{n}}_S=\hat{\bm{n}}_3$. But that is merely the y-component of the capillary vector. The force balance can thus be written as
\begin{equation}\label{eq_youngs_eq_aniso}
     \sigma_{3}(\pi/2)-\sigma_1(-\pi/2)  
       = \sigma_2(\theta_2)\sin(\theta_2) + \sigma_2'(\theta_2)\cos(\theta_2) \,.
 \end{equation}

Geometric interpretation of this equation is that the force equilibrium in the triple junction is achieved when normal angle $\theta_2$ of the particle-parent phase interface in the triple junction is that one, which has the y coordinate of the Wulff shape equal to the difference $\sigma_{3}(\pi/2)-\sigma_1(-\pi/2) $. The equilibrium shape of anisotropic particle on a substrate is thus that of an isolated particle but truncated at a position, which derives from the force balance in the triple junction. The truncated part of isolated particle counts, which is above the truncating line.

When all the three interfaces are isotropic, the well known Young's equation is obtained
\begin{align}
     \sigma_3-\sigma_1 &= \sigma_2\sin(\theta_2) \\
         &= \sigma_2\cos(\vartheta_2)
     \,,
 \end{align}
where in the second equation it was used that $\vartheta_2=\theta_2-\pi/2$.

Let's assume the anisotropy of the particle-parent phase interface $\sigma_2(\theta_2)=\sigma_2^0f(\theta_2)$. Then the wetting parameter $\Gamma$ is defined
\begin{equation}
     \Gamma = \frac{\sigma_{3}(\pi/2)-\sigma_1(-\pi/2) }{\sigma_2^0} \,,
 \end{equation}
which is the vertical offset of the horizontal truncating line relative to center of the equilibrium shape of unit radius. Apparently, when $\Gamma > 0$, the line is above the Wulff shape center (the shape is submerged below the line) and there is such value of $\Gamma$, when the line is merely a tangent. That corresponds to the case of complete wetting and the contact angle is 0. For isotropic interface energy that occurs for $\Gamma=1$.

On the other hand, when $\Gamma < 0$, the truncating line goes below the center (the shape is emerged above the line). There is such value of $\Gamma$ corresponding to complete non-wetting (with isotropic interface energy that is $\Gamma=-1$).

\section{Stability of the equilibrium shape}\label{sec_fund_anisoIE_trijun_stability}
Previous section showed derivation of the 2D equilibrium shape of anisotropic particle on a substrate from the triple junction force balance. However, as explained in~\ref{sec_intro_trijun_forcebalance_stability}, the stability of the solution should be assessed as well. The additional conditions~\eqref{eq_3jun_aniso_stabcond1}, \eqref{eq_3jun_aniso_stabcond2} can be re-written as
\begin{align}
     0 &< \Sigma_1 s_1 + \Sigma_2 s_2 + \Sigma_3 s_3 \\
     0 &< \Sigma_1\Sigma_2 S_{12} + \Sigma_2\Sigma_3 S_{23} + \Sigma_1\Sigma_3 S_{13} \,,
 \end{align}
where $\Sigma_i = \mathrm{d}^2 \sigma_i/\mathrm{d} \vartheta_i^2 + \sigma_i(\vartheta_i)$ are the respective interface stiffnesses, $s_i = \sin^2(\vartheta_i)/d_i$, $S_{ij} = \sin^2(\vartheta_i-\vartheta_j)/d_i d_j$. Given $\vartheta_1=0$ and $\vartheta_3=\pi$, it follows that $s_1=s_3=0$ and also that $S_{13}=0$. Note also that $s_2>0$ and $S_{12}=S_{23}=S=\sin^2(\vartheta_2)/d^2>0$ (assuming that $d_1=d_2=d_3$). The conditions are thus simplified to
\begin{align}
     0 &< \Sigma_2  \label{eq_3jun_stabcond_onplane1}\\
     0 &< \Sigma_1\Sigma_2 + \Sigma_2\Sigma_3 \,,  
 \end{align}
where only the interface stiffnesses matter. Using the first equation, the second may further be simplified
\begin{equation}
     0< \Sigma_1 + \Sigma_3 \,. \label{eq_3jun_stabcond_onplane2}
 \end{equation}
The condition~\eqref{eq_3jun_stabcond_onplane1} implies, that the particle-parent phase interface must be oriented under allowed angle in the contact point. That is fulfilled automatically by the truncated Wulff shape solution, because it contains only the allowed angles. 

The condition~\eqref{eq_3jun_stabcond_onplane2} is surely fulfilled when both interfaces 1 and 3 have positive interface stiffness in their orientations. That holds automatically, when the Wulff shapes of the interfaces 1 and 3 do not contain any corners. When both of them do contain corners, it means that there are normal angles where the interface stiffness is negative. Apparently, in such situation, the condition~~\eqref{eq_3jun_stabcond_onplane2} may be violated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Keep the following \cleardoublepage at the end of this file, 
% otherwise \includeonly includes empty pages.
\cleardoublepage

% vim: tw=70 nocindent expandtab foldmethod=marker foldmarker={{{}{,}{}}}
